Corotating Dust Clumps Around Adolescent M-Dwarfs: Four Years of TESS

### 2023-06-27 paper/ms.tex
update ms

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show quasiperiodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle; sharp peaks
are also common.  Most CQVs are pre-main-sequence M-dwarfs without
near-infrared excesses, with ages of $\approx$5-150 million years
(Myr), and rotation periods of at most two days; they are observed to
be $\approx$1-3\% of M-dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  The dips can be
chromatic, with a reddening law plausibly consistent with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{lc_mosaic_dlt150_good_all.pdf}
		\caption{
			{\bf CQVs from a search of the TESS 2-minute data at
				$d$$<$150\,pc, acquired between July~2018 and Sep~2022.}
			Phased TESS light curves over 1 month are shown for 40 CQVs;
			they include the brightest and closest examples of CQVs known
			($V$=14; $J$=9.5; $d$=25\,pc).  Gray are raw 2-minute data;
			black bins to 300 points per cycle.  Periods in hours are listed
			in the lower right corners of each panel.
			{\bf todo:  add TIC IDs and sector numbers}.
			% TODO: sort by brightness maybe?
			% TODO: sort by period?
		}
		\label{fig:singlemosaic}
	\end{center}
\end{figure*}

@@

\subsection{CQV catalog}

Table~\ref{tab:thetable} lists the \ncpvsfound\ CQVs identified by our
search, and some of their important properties.  \ngoods\ objects
demonstrate what we view as unambiguous characteristics of the CQV
phenomenon in at least one TESS sector, and we refer to these as the
``gold sample''.  The remaining \nmaybes\ CQV candidates were
ambiguous (in our view), and their variability might be caused by say,
multiple starspot groups.  Additional TESS data could help resolve
their classification.

The mosaic in Figure~\ref{fig:singlemosaic} shows phased light curves
for the 40 brightest objects.  The TESS sectors in which these objects
were identified show anywhere from one to eight local minima per
cycle.  Some stars show relatively ordinary modulation during one
continuous portion of the phased light curve, and highly structured
modulation in the remainder of the cycle (e.g. TIC~206544316,
TIC~224283342, TIC~402980664).  Others show structured modulation over
the entire span of a cycle (e.g. TIC~2234692, TIC~401789285,
TIC~425933644, TIC~142173958).  Others still show some mix between
these two modes.

@@

\subsection{Infrared excesses}

@@

{\it The light curve}---
Figure~\ref{fig:lplc} shows an alternative view of the LP-502 light
curve from Figure~\ref{fig:lp}, 
split into ``time chunks'' wherever there was more than half a day of
missing data in the TESS light curve.
We binned the light curve to 15-minute intervals for visual clarity,
and required at least one (120-second cadence) flux measurement per
bin.
Points more than 2.5$\sigma$ above the median are drawn in gray, also
for visual clarity.
Missing data are not drawn (a common plotting error when ``drawing''
light curves, rather than showing a scatter plot).

{\it State switches}---
Figure~\ref{fig:lpzoom} shows an alternative view of cycles 1233-1247.

@@

%
%\begin{figure*}[!t]
%	\begin{center}
%		\centering
%		\includegraphics[width=0.96\textwidth]{TIC4029_zoom.png}
%		\vspace{-0.45cm}
%		\caption{
%			{\bf LP 12-502 cycles 1233-1247}, showing the most dramatic state
%			change observed over the three year baseline.
%		}
%		\label{fig:lpzoom}
%	\end{center}
%\end{figure*}

>>>>

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show quasiperiodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle; sharp peaks
are common, but .  Most CQVs are pre-main-sequence M-dwarfs without
near-infrared excesses, with ages of $\approx$5-150 million years
(Myr), and rotation periods of at most two days; they are observed to
be $\approx$1-3\% of M-dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  The dips can be
chromatic, with a reddening law plausibly consistent with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f2.pdf}
    \vspace{-0.7cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data at
      $d$$<$150\,pc, acquired between July~2018 and Sep~2022.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs.
      Gray are raw 2-minute data;
			black bins to 300 points per cycle.  Each panel is labelled by
      the TIC identifier, the TESS sector number, and the period in hours.
      The ordering is discussed in Section~\ref{sec:catalog}
		}
		\label{fig:singlemosaic}
	\end{center}
\end{figure*}

@@

\subsection{CQV catalog}
\label{sec:catalog}

Table~\ref{tab:thetable} lists the \ncpvsfound\ CQVs identified by our
search, along important physical and observational properties.
\ngoods\ objects demonstrate what we view as unambiguous
characteristics of the CQV phenomenon in at least one TESS sector, and
we refer to these as the ``gold sample''.  The remaining \nmaybes\ CQV
candidates were ambiguous (in our view), and their variability might
be caused by say, multiple starspot groups.  Additional data from TESS
or other instruments can help resolve their classification; in
Section~\ref{subsec:irexcess} below for instance we discuss
a few cases in this ambiguous sample that are probably dippers.

The mosaic in Figure~\ref{fig:singlemosaic} shows phased light curves
for the \ngoods\ bonafide CQVs.  
The objects are sorted first in order of the number of TESS sectors in
which they clearly demonstrated the CQV phenomenon, and secondarily by
descending brightness.
The objects at the top therefore have the most data.

Generally, Figure~\ref{fig:singlemosaic} shows a broad range of
variability, with anywhere from two to eight local minima per cycle.  
Some stars show relatively ordinary modulation during one
continuous portion of the phased light curve, and highly structured
modulation in the remainder of the cycle (e.g. TIC~206544316,
TIC~224283342, TIC~402980664).  Others show structured modulation over
the entire span of a cycle (e.g. TIC~2234692, TIC~401789285,
TIC~425933644, TIC~142173958).  Others still show some mix between
these two modes.

A few objets at first glance seem reminiscent of eclipsing binaries, for
instance TIC~193831684.
In every such case,

@@

\subsection{Infrared excesses}
\label{subsec:irexcess}

@@

{\it The light curve}---
Figure~\ref{fig:lplc} shows an alternative view of the LP-502 light
curve from Figure~\ref{fig:lp}, split into successive TESS orbits.  We
binned the light curve to 15-minute intervals for visual clarity, and
required at least one (120-second cadence) flux measurement per bin.
Points more than 2.5$\sigma$ above the median are drawn in gray, also
for visual clarity.  Missing data are not drawn (a common plotting
error when ``drawing'' light curves, rather than showing a scatter
plot).  The red lines point to specific state switches, discussed
elsewhere in the text.

Figure~\ref{fig:lp2} shows another alternative view of
Figure~\ref{fig:lp}, but arranged to enable easy visual appreciation
of transit timing changes, rather than transit depth changes.
A best-fitting two-harmonic sinusoid has been independently fitted and
subtracted from the Sector
18-19 data, 25-26 data, and 53, 58, and 59 data.

Finally, Figure~\ref{fig:lpriver0} shows ``river plots'' of the
same data, split into very similar intervals:
the Sector 18-19 data, 25-26 data, 53 data, and 58-59 data.
State changes are evident in these plots whenever there is a sudden
change in color.
{\bf todo: fix missing data to be different from the flare color}

@@
